---
title: "HCA"
author: "Justin Murray"
date: "10/20/2019"
output: html_document
---
## Hierarchical clustering
Hierarchical clustering (HC) creates a hierarchy of clusters and presents them in a dendrogram. HC does not reuire the user to specify a number of clusters to begin. HC is a technique that tries to but a cluster with two steps: 

1. Agglomerative - bottom up
2. Divisive - top down

Before performing HC we need to determine how to figure out the distance to create clusters of sameness. There are Four distance functions to measusre similarity

1. Single Linkage - Shortest distance between two points in a cluster
2. Complete Linkage - Longest distance between two points in a cluster
3. Average Linkage - Average distance between two points in a cluster
4. Ward method - Sum of the Squared distance from each point to the mean of the merged clusters


```{r include=FALSE}
rm(list = ls()) 
graphics.off()
cat("\014") 

library("dendextend")
library("magrittr")

# source("C:\\Users\\pandr\\OneDrive\\R_code\\installPackages.R")
setwd('/Users/justinmurray/Desktop/ML680/Week6/data')
whole = read.csv("Wholesale_customers_data.csv")
whole = scale(whole, scale = TRUE)
hc = hclust(dist(whole, method="euclidean"), method = "ward.D2")
```

## Ward Linkage

```{r echo=FALSE}
plot(hc, hang = -0.01, cex = 0.7, labels = FALSE)
```

##Single Linkage
```{r echo=FALSE}
hc2 = hclust(dist(whole), method="single")
plot(hc2, hang = 0.1, cex = 0.7, labels = FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# divisive Hierarchiacl clustering
hcd = as.dendrogram(hc)
plot(hcd, xlab = "Height", horiz = TRUE, labels = FALSE)
hcd %>% set("branches_k_color", k = 3) %>% plot(horiz = TRUE)
hcd %>% rect.dendrogram(k = 3, horiz = TRUE, border = 8, lty = 5, lwd = 2 , labels=FALSE)
abline(v = heights_per_k.dendrogram(hcd)["4"] + .1, lwd = 2, lty = 2, col = "blue")
```